---
title: "Math 5810 - Homework 5"
author: "Professor X"
date: "`r strftime(Sys.Date(), format = '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

# Introduction

Using Hadley's `bigvis` for the airline data 2013 to 2016

To use this code, first run the code in install-bigvis.R then unzip the file Airline.7z (I use 7-zip) into the folder `../Data/Airline`. You can adjust the folder relative to your desired working directory. Usually I assume the working directory is the home of `bigvis-06.R`

# Setup

```{r}
devtools::session_info()
```

```{r, message=FALSE}
library(magrittr)   # pipe operators
library(data.table) # fast/effiecient read
library(tidyverse)  # loads dplyr, ggplot2, reshape2, etc.
                    # https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
library(bigvis)
library(scales)
# library(grid)
```

# Get the Data

`../Data/Airline` contains lots of data files. We can read the file names:

```{r}
fnames <- list.files(
    path       = '../../data/airline-cut'
  , full.names = T
  )
```

You should check that fnames contains the file names. Otherwise you have a problem with where you stored your data or the path name above

```{r}
fnames
```

Read in the contents of each file and combine them, then free up a bunch of memory by reassigning the initial list object we used to read in the data:

```{r}
flights <- lapply(
    X          = fnames
  , FUN        = data.table::fread
  , data.table = F
  )

flights %<>%
  bind_rows() %>% 
  as_data_frame()
```

# Subset the data

The variables we will use are the following:

```{r}
flights %<>%
  select(
      DAY_OF_WEEK
    , DEP_TIME
    , ARR_TIME
    , ARR_DELAY
    , DEP_DELAY
    , AIR_TIME
    , DISTANCE
  )
```

Notice that some of the values are missing. You can remove them if you like using
`flights = na.omit(flights)`. I'm using the original with missing values included:

```{r}
flights %>% 
  is.na() %>% 
  sum()
```

Rename some columns for convenience:

```{r}
names(flights) <- c(
    'dayofweek'
  , 'deptime'
  , 'time'
  , 'delay'
  , 'depdelay'
  , 'airtime'
  , 'dist'
)
```

Add our calculated fields:
```{r}
flights %<>%
  mutate(
      speed = dist / (airtime / 60)
    , hours = floor(time / 100)
    , mins  = time - 100 * hours
    , time  = hours + (mins / 60)
  )
```

Time is in minutes, speed in miles per hour.

# Viz Creation

## Day with most flights

This is where I was going to find out if there was a day of the week that had more flights. If I did this right there are more flights on Monday than the rest of the week.

```{r, message=FALSE}
dayofweek_sum <- bin(
    x     = flights$dayofweek
  , width = 1
  ) %>% 
  condense()

dayofweek_sum %>%
  autoplot()

dayofweek_sum %>% 
  smooth(h = 10) %>% 
  autoplot()
```


## Departure delays

This one is trying to show how many departure delays across the data and it seems like most made it out on time or earlier.

```{r, message=FALSE}
depdelay_sum <- bin(
    x     = flights$depdelay
  , width = 100
  ) %>% 
  condense()

depdelay_sum %>% 
  autoplot()

depdelay_sum %>% 
  smooth(h = 1000) %>% 
  autoplot()
```

## On time departures by day of week

I tried to see if there day of week made a difference of departing on time. I think it also shows that most flights leave in the morning and all the later flights happen during midweek.

```{r, message=FALSE}
daydelay <- condense(
    bin(flights$dayofweek, 1)
  , bin(flights$depdelay, 60)
  )

daydelay %>% 
  autoplot()

daydelay %>% 
  peel() %>% 
  autoplot()
```

## Airtime

```{r, message=FALSE}
teaser <- list(
    theme(
      legend.position  = "bottom"
    , plot.margin      = unit(c(0, 0.5, 0, 0.5), "lines")
    , legend.key.width = unit(1.45, "inches")
    , text             = element_text(size = 24)
  )
  , labs(
      x    = NULL
    , y    = NULL
    , fill = NULL
    )
)

daydeld <- condense(
    bin(flights$dayofweek, 1)
  , bin(flights$depdelay, 120)
  , z = flights$airtime
  )

daydeld %<>%
  subset(flights.dayofweek > 0)

daydeld %>% 
  autoplot() + teaser


airtime_sum <- bin(
    x     = flights$airtime
  , width = 100
  ) %>% 
  condense()

airtime_sum %>% 
  autoplot()

airtime_sum %>% 
  smooth(h = 1000) %>% 
  autoplot()
```

## Airtime vs Departure delays

This one was comparing airtime and departure delay. It looks to me like the more airtime you have on a flight, the less likely you are to depart late.

```{r, message=FALSE}
airdelay <- condense(
    bin(flights$airtime, 10)
  , bin(flights$depdelay, 10)
  )

airdelay %>% 
  autoplot()

airdelay %>% 
  peel() %>% 
  autoplot()

airdelayd <- condense(
    bin(flights$airtime, 10)
  , bin(flights$depdelay, 10)
  , z = flights$dist
  )
airdelayd %<>%
  subset(flights.airtime > 100)

airdelayd %>% 
  autoplot() + teaser
```
